Summary Tables

### 2023-12-06 summary_tables.Rmd
Uncomment code to generate health outcomes as CSV files

---
title: "Summary Tables"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  output_version: ''
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: true
---

@@

<!-- ```{r, echo=FALSE, results='hide', message=FALSE} -->

<!-- # Export all health results -->

<!-- # hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results -->
<!-- # pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA) -->

<!-- # initialise variables -->
<!-- health <- list() -->
<!-- measure <- c("ylls", "deaths") -->
<!-- pathway <- c("hb", "pathway_hb") -->

<!-- for (k in measure) { # loop for ylls or deaths -->
<!--   for (j in pathway) { # loop for hb or pathway_hb -->
<!--     for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities -->

<!--       # Temporal dataset for health outcomes -->
<!--       temp <- io[[i]]$outcomes[[j]][[k]] -->

<!--       # Total population -->
<!--       #overall_pop <- sum(io[[i]]$demographic$population) -->

<!--       # Population by age and sex -->
<!--       pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>%  -->
<!--         summarize(pop_age_sex = sum(population)) -->

<!--       # get the health burden into the correct format by creating a long format and adding additional columns -->
<!--       # create long format -->
<!--       health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure",  -->
<!--                                       -sex, -age_cat) %>%  -->
<!--         mutate(city = i, # create city column -->
<!--                scenario = case_when( # create scenario column -->
<!--                  grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'), -->
<!--                  grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'), -->
<!--                  grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'), -->
<!--                  grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle') -->
<!--                ), -->
<!--                level1 = case_when( # create level 1 column -->
<!--                  grepl("all_cause", cause) ~ "L1: All Cause", -->
<!--                  grepl("inj", cause) ~ "L1: RTI" -->
<!--                ), -->
<!--                level2 = case_when( # create level 2 column -->
<!--                  grepl("total_cancer", cause) ~ "L2: Cancer", -->
<!--                  grepl("CVD", cause) ~ "L2: CVD", -->
<!--                  grepl("respiratory", cause) ~ "L2: Respiratory", -->
<!--                  grepl("T2D", cause) ~ "L2: Other", -->
<!--                  grepl("total_dementia", cause) ~ "L2: Other", -->
<!--                  grepl("Parkinson", cause) ~ "L2: Other", -->
<!--                  grepl("inj", cause) ~ "L2: RTI" -->
<!--                ), -->
<!--                level3 = case_when( # create level 3 column -->
<!--                  grepl("IHD", cause) ~ "L3: IHD", -->
<!--                  grepl("lung_cancer", cause) ~ "L3: lung_cancer", -->
<!--                  grepl("COPD", cause) ~ "L3: COPD", -->
<!--                  grepl("stroke", cause) ~ "L3: stroke", -->
<!--                  grepl("T2D", cause) ~ "L3: T2D", -->
<!--                  grepl("LRI", cause) ~ "L3: LRI", -->
<!--                  grepl("breast_cancer", cause) ~ "L3: breast_cancer", -->
<!--                  grepl("colon_cancer", cause) ~ "L3: colon_cancer", -->
<!--                  grepl("endo_cancer", cause) ~ "L3: endo_cancer", -->
<!--                  grepl("liver_cancer", cause) ~ "L3: liver_cancer", -->
<!--                  grepl("total_dementia", cause) ~ "L3: total_dementia", -->
<!--                  grepl("myeloma", cause) ~ "L3: myeloma", -->
<!--                  grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia", -->
<!--                  grepl("Parkinson", cause) ~ "L3: Parkinson", -->
<!--                  grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer", -->
<!--                  grepl("stomach_cancer", cause) ~ "L3: stomach_cancer", -->
<!--                  grepl("inj", cause) ~ "L3: RTI" -->
<!--                ), -->
<!--                dose = case_when( # create dose column -->
<!--                  grepl("_pa_ap_", cause) ~ "PA and AP", -->
<!--                  grepl("_pa_", cause) ~ "PA", -->
<!--                  grepl("_ap_", cause) ~ "AP", -->
<!--                  grepl("_inj", cause) ~ "RTI" -->
<!--                ) -->
<!--         ) %>% # End mutate  -->
<!--         # join with demographic information -->
<!--         left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))  -->

<!--         # %>% -->
<!--         # create new columns showing burden per 100000 people by age or by age and sex -->
<!--         # mutate(metric_100k = measure / overall_pop * 100000,  -->
<!--         #        metric_100k_sex = measure / overall_pop * 100000) -->

<!--     } # Loop for each city -->
<!--   } # Loop for each pathway -->
<!-- } # Loop for each measure -->

<!-- # for each health burden and pathway join all cities into one dataframe -->
<!-- ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
<!-- ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
<!-- deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
<!-- deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->

<!-- # # Export ylls HB -->
<!-- # output_file <- "results/multi_city/health_impacts/data.xlsx" -->
<!-- # writexl::write_xlsx(list(ylls = ylls, -->
<!-- #                          ylls_pathway = ylls_pathway, -->
<!-- #                          deaths = deaths, -->
<!-- #                          deaths_pathway = deaths_pathway),  -->
<!-- #                     path = output_file) -->

<!-- # Export ylls HB -->

<!-- # write csv with output version number -->
<!-- write.csv(ylls, -->
<!--             paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'), -->
<!--             row.names = F) -->
<!-- # Export ylls PATHWAY_HB -->
<!-- write.csv(ylls_pathway, -->
<!--             paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'), -->
<!--             row.names = F) -->
<!-- # Export deaths HB -->
<!-- write.csv(deaths, -->
<!--             paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'), -->
<!--             row.names = F) -->
<!-- # Export deaths PATHWAY_HB -->
<!-- write.csv(deaths_pathway, -->
<!--             paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'), -->
<!--             row.names = F) -->

<!-- # write csv without output version number -->
<!-- write.csv(ylls, -->
<!--             paste0('results/multi_city/health_impacts/ylls.csv'), -->
<!--             row.names = F) -->
<!-- # Export ylls PATHWAY_HB -->
<!-- write.csv(ylls_pathway, -->
<!--             paste0('results/multi_city/health_impacts/ylls_pathway.csv'), -->
<!--             row.names = F) -->
<!-- # Export deaths HB -->
<!-- write.csv(deaths, -->
<!--             paste0('results/multi_city/health_impacts/deaths.csv'), -->
<!--             row.names = F) -->
<!-- # Export deaths PATHWAY_HB -->
<!-- write.csv(deaths_pathway, -->
<!--             paste0('results/multi_city/health_impacts/deaths_pathway.csv'), -->
<!--             row.names = F) -->

>>>>

---
title: "Summary Tables"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  output_version: ''
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: 
      collapsed: false
    number_sections: true
---

@@

```{r, echo=FALSE, results='hide', message=FALSE}

# Export all health results

# hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results
# pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA)

# initialise variables
health <- list()
measure <- c("ylls", "deaths")
pathway <- c("hb", "pathway_hb")

for (k in measure) { # loop for ylls or deaths
  for (j in pathway) { # loop for hb or pathway_hb
    for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities

      # Temporal dataset for health outcomes
      temp <- io[[i]]$outcomes[[j]][[k]]

      # Total population
      #overall_pop <- sum(io[[i]]$demographic$population)

      # Population by age and sex
      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>%
        summarize(pop_age_sex = sum(population))

      # get the health burden into the correct format by creating a long format and adding additional columns
      # create long format
      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure",
                                      -sex, -age_cat) %>%
        mutate(city = i, # create city column
               scenario = case_when( # create scenario column
                 grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'),
                 grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'),
                 grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'),
                 grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle')
               ),
               level1 = case_when( # create level 1 column
                 grepl("all_cause", cause) ~ "L1: All Cause",
                 grepl("inj", cause) ~ "L1: RTI"
               ),
               level2 = case_when( # create level 2 column
                 grepl("total_cancer", cause) ~ "L2: Cancer",
                 grepl("CVD", cause) ~ "L2: CVD",
                 grepl("respiratory", cause) ~ "L2: Respiratory",
                 grepl("T2D", cause) ~ "L2: Other",
                 grepl("total_dementia", cause) ~ "L2: Other",
                 grepl("Parkinson", cause) ~ "L2: Other",
                 grepl("inj", cause) ~ "L2: RTI"
               ),
               level3 = case_when( # create level 3 column
                 grepl("IHD", cause) ~ "L3: IHD",
                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
                 grepl("COPD", cause) ~ "L3: COPD",
                 grepl("stroke", cause) ~ "L3: stroke",
                 grepl("T2D", cause) ~ "L3: T2D",
                 grepl("LRI", cause) ~ "L3: LRI",
                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
                 grepl("total_dementia", cause) ~ "L3: total_dementia",
                 grepl("myeloma", cause) ~ "L3: myeloma",
                 grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia",
                 grepl("Parkinson", cause) ~ "L3: Parkinson",
                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
                 grepl("inj", cause) ~ "L3: RTI"
               ),
               dose = case_when( # create dose column
                 grepl("_pa_ap_", cause) ~ "PA and AP",
                 grepl("_pa_", cause) ~ "PA",
                 grepl("_ap_", cause) ~ "AP",
                 grepl("_inj", cause) ~ "RTI"
               )
        ) %>% # End mutate
        # join with demographic information
        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))

        # %>%
        # create new columns showing burden per 100000 people by age or by age and sex
        # mutate(metric_100k = measure / overall_pop * 100000,
        #        metric_100k_sex = measure / overall_pop * 100000)

    } # Loop for each city
  } # Loop for each pathway
} # Loop for each measure

# for each health burden and pathway join all cities into one dataframe
ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))

# # Export ylls HB
# output_file <- "results/multi_city/health_impacts/data.xlsx"
# writexl::write_xlsx(list(ylls = ylls,
#                          ylls_pathway = ylls_pathway,
#                          deaths = deaths,
#                          deaths_pathway = deaths_pathway),
#                     path = output_file)

# Export ylls HB

# write csv with output version number
write.csv(ylls,
            paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'),
            row.names = F)
# Export ylls PATHWAY_HB
write.csv(ylls_pathway,
            paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'),
            row.names = F)
# Export deaths HB
write.csv(deaths,
            paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'),
            row.names = F)
# Export deaths PATHWAY_HB
write.csv(deaths_pathway,
            paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'),
            row.names = F)

# write csv without output version number
write.csv(ylls,
            paste0('results/multi_city/health_impacts/ylls.csv'),
            row.names = F)
# Export ylls PATHWAY_HB
write.csv(ylls_pathway,
            paste0('results/multi_city/health_impacts/ylls_pathway.csv'),
            row.names = F)
# Export deaths HB
write.csv(deaths,
            paste0('results/multi_city/health_impacts/deaths.csv'),
            row.names = F)
# Export deaths PATHWAY_HB
write.csv(deaths_pathway,
            paste0('results/multi_city/health_impacts/deaths_pathway.csv'),
            row.names = F)
